# Logging

::: callout-note
## Key Points

-   Version control your RAP using Git

-   Use Github to collaborate on your RAP

-   Do not push your data or sensitive parameters to Github

-   Keep an up to date version of your RAP and your data on Sharepoint.
:::

If we are going to batch run scripts, or set them to run automatically, we need a way of recording what they are doing, whether they have run correctly, and any issues that have occurred. To do this we use logging.

Logging is simply adding lines to your code that record outputs into a log file, stored in our project folders.

We will look at:

-   Setting up logger
-   Basic Structure of using logger

## Set up

To set up logger, the following code is used:

```{r, eval = F}
if (!requireNamespace("librarian")) install.packages("librarian")
librarian::shelf(DataS-DHSC/DHSClogger)
```

Here we are using DHSClogger, a package written to simplfy the logging process.

R has three levels of message it can return in your console: `INFO`, `WARN` and `ERROR`. In the above code we we set our threshold to `INFO`, meaning all messages `INFO` and up are logged. If this is changed to i.e. `ERROR`, only error mesages will be logged.

The DHClogger is a wrapper for a package for a package called `futile.logger`. You can find more information [here](https://github.com/zatonovo/futile.logger).

## Basic structure

To add something to the log, we use `logger$info()`. Here is the general structure:

```{r, eval = F}
#Start the logger
logger$info("[Begin]")

# add source of run script and entry point to code below
source("./R/main.R", local = TRUE)
run_analysis()

#Stop the logger
logger$info("[End]")
```

To have the save a log message to the log.txt file, we use `logger$info(log_message)`. This will add the log_message variable to the logfile with a time stamp so you can see what result was produced and when.

```{r, eval = F}
rowcount <- nrow(df_data)
logger$info(sprintf("Number of rows downloaded for %s was: %d",
                       config$data_id,
                       rowcount))
```

::: callout-note
## Exercise:

1.  Set up your logger

2.  Remember to add logging in subsequent steps
:::
