# Downloading data {#sec-download}

::: callout-note
## Key Points

-   Be clear about where your data is coming from

-   Wrap up the ingest of your data into a function
:::

Once you have set up your wrap project and got your documentation and logging in place, you can put together your analysis scripts. As discussed in chapter 02, you should make your code as modular as possible, wrapping different processes in functions, then sourcing and calling them from a central script.

The data for your wraps can come from a variety of sources. Ideally you want this data to be available in the same format each time you run your wrap, to avoid needing to rewrite. This can be from published sources, from a database, or the product of other analysis.

For the example in this book we will use data from the ONS api.

## Data setup and loading ONS meta data

We start by adding the packages we will be using to the `main.R` script. Keeping the reading of your packages central means you only need to do it once.

We can use the librarian package to manage installation and reading of our packages. It handles the installation of packages from a variety of sources, so if you are using a packages from Github as well as CRAN, the process is streamlined.

The below code adds the packages we will need for the data ingest process. It loads `httr` and `jsonlite` for interacting with the api and it's response, `dplyr` for manipulating data, and `DHSClogger` from the DHSC Data Science Github for logging.

```{r, eval = F}
if (!requireNamespace("librarian")) install.packages("librarian")
librarian::shelf(readr, httr, jsonlite, dplyr, DataS-DHSC/DHSClogger)
```

It is best practice to define parameters needed in your code in configuration files outside of the scripts themselves. This means you don't have to touch the code when you need to change any parameter values.

In the DHSC Template, the main configuration file sits in the ./input directory and is called `config.yml`. A YAML is a human readable data serialization language, it is a useful method for storing parameters in an easily understandable way.

The below parameters are the ones we will use to access data from the ONS api.

    data_dictionary_url: "https://api.beta.ons.gov.uk/v1/datasets"
    data_id: "regional-gdp-by-year"
    input_dir: ./input
    output_dir: ./output

For our script to use the values stored in `config.yml` we need to read it in. This is done in the `main.R` script.

```{r, eval = F}
# read in configuration
config_path <- file.path("input", "config.yml")
config <- yaml::read_yaml(config_path)

# print the config file details
print(config)
```

## Get a list of data that can be extracted using this ONS API endpoint

We pass the data_dictionary_url to a GET function to download a list of available data.

```{r, eval = F}
# request the data dictionary from the ONS api endpoint
path <- config$data_dictionary_url

request <- httr::GET(url = path)
      
content <- rawToChar(request$content)
      
content_flat <- json.lite::fromJSON(content, flatten = TRUE) 
      
logger$info(paste0("Response recieved from: ", path))
      
# create a dataframe with the downloaded data dictionary to what data is available
df <- data.frame(id = content_flat$items$id,
                 title = content_flat$items$title,
                 release_frequency = content_flat$items$release_frequency,
                 next_release = content_flat$items$next_release,
                 description = content_flat$items$description,
                 links_latest_versions = content_flat$items$links.latest_version.href,
                 links_self_href= content_flat$items$links.self.href,
                 qmi_href= content_flat$items$qmi.href)
```

The data dictionary contents are shown below

```{r, echo = FALSE, warning = FALSE}
library(readr)
library(htmlTable)
htmlTable(read_csv("input/data_dictionary_response.csv"), rnames = FALSE)

```

## Get the URL for the latest release of a specified ONS file

ONS data files have a unique identifier and this can be used to extract the data. The id we are interested in is specified in the config file and so use that to interrogate the data dictionary. In particular, we want to extract the latest version, the url for which can be extracted from the data dictionary in the links_latest_versions field we have just created.

```{r, eval = F}
file_latest_version <- df %>%
  dplyr::filter(id == config$data_id) %>%
  dplyr::select(links_latest_versions) %>%
  dplyr::pull()

# Using the links_latest_versions we can get the meta data for the latest file
# and this meta data includes the url for the latest file in either csv or xlsx
# format. We choose .csv here. Extract it from downloads$csv$href

request <- httr::GET(url = file_latest_version)
      
logger$info(paste0("Response recieved from: ", file_latest_version))
      
content <- rawToChar(request$content)
content_flat <- jsonlite::fromJSON(content, flatten=TRUE)
file_url <- content_flat$downloads$csv$href
```

## Now we have the file url we can read it like any other file

```{r, eval = F}
#Read data from latest file endpoint
df_data <- readr::read_csv(file_url)

data_filepath <- paste0(config$output_dir, "/", config$data_id, ".csv")

readr::write_csv(df_data, data_filepath)

logger$info(paste0("data output to: ", data_filepath))
```

## Wrap in function

Now we have all the code we need we can wrap it up in a function to be sourced and called in our `run_analysis.R` script, making sure to add our Roxygen documentation.

```{r, eval = F}
#' Downloads a csv from the ONS API, and saves to input folder
#'
#' @param data_id a string of the name of the data set to download. This is stored in config.yml as config$data_id
#'
#' @import dplyr
#' @import httr
#' @import jsonlite
#' @import logger
#' @import readr
#'
#'
download_data_from_ons <- function(data_id){

    # request the data dictionary from the ONS api endpoint
      path <- config$data_dictionary_url
    
      request <- httr::GET(url = path)
            
      content <- rawToChar(request$content)
          
      content_flat <- jsonlite::fromJSON(content, flatten = TRUE) 
            
      logger$info(paste0("Response recieved from: ", path))
            
        # create a dataframe with the downloaded data dictionary to what data is available
      df <- data.frame(id = content_flat$items$id,
                       title = content_flat$items$title,
                       release_frequency = content_flat$items$release_frequency,
                       next_release = content_flat$items$next_release,
                       description = content_flat$items$description,
                       links_latest_versions = content_flat$items$links.latest_version.href,
                       links_self_href= content_flat$items$links.self.href,
                       qmi_href= content_flat$items$qmi.href)
      
      
 file_latest_version <- df %>%
 dplyr::filter(id == data_id) %>%
 dplyr::select(links_latest_versions) %>%
 dplyr::pull()
      
  # Using the links_latest_versions we can get the meta data for the latest file
  # and this meta data includes the url for the latest file in either csv or xlsx
  # format. We choose .csv here. Extract it from downloads$csv$href
      
  request <- httr::GET(url = file_latest_version)
      
  logger$info(paste0("Response recieved from: ", file_latest_version))
        
  content <- rawToChar(request$content)
  content_flat <- jsonlite::fromJSON(content, flatten=TRUE)
  file_url <- content_flat$downloads$csv$href
      
      
  #Read data from latest file endpoint
  df_data <- readr::read_csv(file_url)
      
  data_filepath <- paste0(config$output_dir, "/", data_id, ".csv")
      
  readr::write_csv(df_data, data_filepath)
      
  logger$info(paste0("data output to: ", data_filepath))
      
}
```

::: callout-note
## Exercise: Ingest your data

1.  Alter the config file to look at a different data set

2.  Run the ingest script to make sure you are reading in your data corrctly
:::
