# Processing Data 

::: callout-note
## Key Points

-   Set out your processing in clear steps

-   Wrap up the processing of your data into a function
:::

# Cleaning and processing your data

This page will take you through the steps for the data analysis we wish to do. It is an example of some simple data cleaning and sorting, showing how you can parameterise and modularise your analysis.

## Reading in the data

For this section we need two packages, `dplyr` and `readr`. We have already included these in `main.R`

First, we can print the column names of the dataset and the unique industry types in the dataset.

```{r, eval = F}
#read in data from data ingest

data_file_path <- paste0(config$output_dir, "/", data_id, ".csv")

data <- read_csv(data_file_path)

logger$info(paste0("Data read from: ", data_file_path))
```

## Filtering data

We will now clean the data and filter for data on the health and social care industry in England and by type of growth rate we wish to analyse. We can parameterise the variables we are interested in and add them to our config file. This will make it straight forward to change later to perform the same analysis on a different industry. The column containing the values we are interested in is called `v4_1`

```{r, eval = F}
# remove NAs, filter by industry, geography and growth rate figure

industry <- config$industry_name
geography <- config$geography
metric <- config$metric

data_time_series <- data %>%
  filter(!is.na(v4_1)) %>%
  filter(UnofficialStandardIndustrialClassification == industry) %>%
  filter(Geography == geography) %>%
  filter(GrowthRate == metric)
      
# set year as a number and sort by
data_time_series_sorted <- data_time_series %>%
  mutate(Time = as.numeric(Time)) %>%
  arrange(Time)
```

Once we have a cleaned dataset, we can save it as a csv for future use.

```{r, eval = F}
# save data to csv

output_file_name <- paste0("output/"config$industry_name, "_", config$metric, "time_series_sorted.csv")

write_csv(
  data_time_series_sorted,
  file = "./output/health_gdp_time_series_sorted.csv")
      
logger$info(paste0("data saved to ", output_file_name))
```

## Wrap in function

As with the initial ingestion of the data, we should wrap our code in a function to be called from our `run_analysis.R` script.

```{r, eval = F}
#' Filters relevant data and sorts it
#'
#' @param data_id a string of the name of the data set to download. This is stored in config.yml as config$data_id
#' @param industry is a string of the industy we want to look at. This is stored in config.yml as config$industry
#' @param geography is a string of the geography we want to look at. This is stored in config.yml as config$geography
#' @param metric is a string of the metric we want to look at. This is stored in config.yml as config$metric
#'
#'
#' @import dplyr
#' @import logger
#' @import readr
#'
#'
process_data <- function(data_id, industry, geography, metric){

  #read in data from data ingest
    
  data_file_path <- paste0(config$output_dir, "/", data_id, ".csv")
      
  data <- readr::read_csv(data_file_path)
      
  logger$info(paste0("Data read from: ", data_file_path))
      
  data_time_series <- data %>%
    filter(!is.na(v4_1)) %>%
    filter(UnofficialStandardIndustrialClassification == industry) %>%
    filter(Geography == geography) %>%
    filter(GrowthRate == metric)
        
  # set year as a number and sort by
  data_time_series_sorted <- data_time_series %>%
    mutate(Time = as.numeric(Time)) %>%
    arrange(Time)
        
  # save data to csv
      
  output_file_name <- paste0("output/", data_id, "_", industry, "_", metric, "_", geography, "_time_series_sorted.csv")
      
  write_csv(
    data_time_series_sorted,
    file = output_file_name)
        
  logger$info(paste0("processed data saved to ", output_file_name))
      
  return(data_time_series_sorted)

}
```

The last steps are:

-   Add new functions to `librarian` line in `main.R`
-   Add a `source` to your new function to `run_analysis.R`
-   Add the function to your function in `run_analysis.R`

::: callout-note
## Exercise: Process your data

1.  Alter the config file to look to process your data correctly

2.  Run the process data script to make sure it's working correctly. Make any necessary fixes.
:::
