# Example scripts

::: {.panel-tabset .nav-pills}
# main.r

```{r, eval=F}
# install packages if needed

if (!requireNamespace("librarian")) install.packages("librarian", quiet = TRUE)

suppressWarnings(
  librarian::shelf(
    DataS-DHSC/DHSClogger,
    ggplot2,
    httr,
    jsonlite,
    dplyr,
    readr,
    DHSCcolours,
    quiet = TRUE))

# Setup logging
logger <- DHSClogger::get_dhsc_logger()

# set threshold of console log to information and above
logger$set_threshold("log.console", "INFO")

# read in config.yml
config <- yaml::read_yaml(
  file.path("input", "config.yml")
)

#run analysis
source("r/run_analysis.r")
run_analysis()


logger$info("[End]")

```

# run_analysis.r

```{r, eval=F}


run_analysis <- function(){

  #read in your analysis functions
  source("r/ingest_data.r")
  source("r/process_data.r")
  source("r/plot_data.r")

  #read in data
  download_data_from_ons(data_id = config$data_id)

  #process data
  ons_data <- process_data(data_id = config$data_id,
                           industry = config$industry,
                           geography = config$geography,
                           metric = config$metric)

  #plot data
  chart <- ts_line_chart_data(data = ons_data,
                              industry = config$industry,
                              metric = config$metric,
                              title = config$chart_title,
                              subtitle = config$subchart_title,
                              xlab = config$xlab,
                              ylab = config$ylab)

  chart

}


```

# config.yml

    # see https://cran.r-project.org/web/packages/ymlthis/vignettes/yaml-overview.html
    # for details of writing yaml files

    data_dictionary_url: "https://api.beta.ons.gov.uk/v1/datasets"
    data_id: "regional-gdp-by-year"
    input_dir: ./input
    output_dir: ./output
    industry : "Q: Human health and social work activities"
    geography : "England"
    metric : "Annual growth rate"
    chart_title : "Annual growth rate of Human health and social work activities, England"
    subtitle : "2013 to 2021"
    xlab : "Year"
    ylab : "Annual growth rate (%)"

# ingest_data.r

```{r, eval=F}
#' Downloads a csv from the ONS API, and saves to input folder
#'
#' @param data_id a string of the name of the data set to download. This is stored in config.yml as config$data_id
#'
#' @import dplyr
#' @import httr2
#' @import jsonlite
#' @import logger
#' @import readr
#'
#'
download_data_from_ons <- function(data_id){

    # request the data dictionary from the ONS api endpoint
      path <- config$data_dictionary_url

  content_flat <- read_json(path, simplifyVector = T)

  logger$info(paste0("Response recieved from: ", path))

  # create a dataframe with the downloaded data dictionary to what data is available
  df <- data.frame(id = content_flat$items$id,
                   title = content_flat$items$title,
                   release_frequency = content_flat$items$release_frequency,
                   next_release = content_flat$items$next_release,
                   description = content_flat$items$description,
                   links_latest_versions = content_flat$items$links$latest_version$href,
                   links_self_href= content_flat$items$links$self$href,
                   qmi_href= content_flat$items$qmi$href)


      
      
 file_latest_version <- df %>%
 filter(id == data_id) %>%
 select(links_latest_versions) %>%
 pull()
      
  # Using the links_latest_versions we can get the meta data for the latest file
  # and this meta data includes the url for the latest file in either csv or xlsx
  # format. We choose .csv here. Extract it from downloads$csv$href
      
  request <- GET(url = file_latest_version)
      
  logger$info(paste0("Response recieved from: ", file_latest_version))
        
  content <- rawToChar(request$content)
  content_flat <- fromJSON(content, flatten=TRUE)
  file_url <- content_flat$downloads$csv$href
      
      
  #Read data from latest file endpoint
  df_data <- read_csv(file_url)
      
  data_filepath <- paste0(config$output_dir, "/", data_id, ".csv")
      
  write_csv(df_data, data_filepath)
      
  logger$info(paste0("data output to: ", data_filepath))
  
  return(data_filepath)
      
}
```

# process_data.r

```{r, eval = F}
#' Filters relevant data and sorts it
#'
#' @param data_id a string of the name of the data set to download. This is stored in config.yml as config$data_id
#' @param industry is a string of the industy we want to look at. This is stored in config.yml as config$industry
#' @param geography is a string of the geography we want to look at. This is stored in config.yml as config$geography
#' @param metric is a string of the metric we want to look at. This is stored in config.yml as config$metric
#'
#'
#' @import dplyr
#' @import logger
#' @import readr
#'
#'
process_data <- function(data_id, industry, geography, metric){

  #read in data from data ingest
    
  data_file_path <- paste0(config$output_dir, "/", data_id, ".csv")
      
  data <- readr::read_csv(data_file_path)
      
  logger$info(paste0("Data read from: ", data_file_path))
      
  data_time_series <- data %>%
    filter(!is.na(v4_1)) %>%
    filter(UnofficialStandardIndustrialClassification == industry) %>%
    filter(Geography == geography) %>%
    filter(GrowthRate == metric)
        
  # set year as a number and sort by
  data_time_series_sorted <- data_time_series %>%
    mutate(Time = as.numeric(Time)) %>%
    arrange(Time)
        
  # save data to csv
      
  output_file_name <- paste0("output/", data_id, "_", industry, "_", metric, "_", geography, "_time_series_sorted.csv")
      
  write_csv(
    data_time_series_sorted,
    file = output_file_name)
        
  logger$info(paste0("processed data saved to ", output_file_name))
      
  return(data_time_series_sorted)

}

```

# plot_data.r

```{r, eval = F}
#' Plot data as time series
#'
#' @param data_id a string of the name of the data set to download. This is stored in config.yml as config$data_id
#' @param industry is a string of the industy we want to look at. This is stored in config.yml as config$industry
#' @param metric is a string of the metric we want to look at. This is stored in config.yml as config$metric
#' @param title is a string that will be the title of the chart
#' @param subtitle is a string that will be the subtitle of the chart
#' @param xlab is a string that is the x axis label
#' @param ylab is a string that is the y axis label
#'
#' @import DHSCtools
#' @import ggplot2
#'
#'

ts_line_chart_data <- function(data, industry, metric, title, subtitle, xlab, ylab){

  graph <- ggplot(data = data_time_series_sorted,
         aes(Time,
             metric,
             industry)) +
    DHSCcolours::theme_dhsc() +
    geom_line(linewidth = 1) +
    geom_point(size = 3) +
    theme(legend.position="none") +
    DHSCcolours::scale_colour_dhsc_d() +
    labs(
      title = title,
      subtitle = subtitle,
      x = xlab,
      y = ylab) +
    scale_x_continuous(breaks=seq(2013,2021,1))
        
        
  #save chart
  chart_path <- paste0("./output/", metric, "_", industry, "_chart.svg")
    
  ggsave(chart_path,
         plot = graph,
         height = 5,
         width = 10,
         units="in",
         dpi=300)
      
  logger$info(paste0("Graph saved to: ", chart_path))
               return(graph)
      
}

```
:::
